---
title: "MB recreate Table1"
author: "Mara Bodenmann"
date: today
toc: true
abstract-title: "Research question:"
abstract: ""
format:
  html:
    embed-resources: true #single HTML file with no supportive directory
#    css: ../../styles.css
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false #this prevent warning messages being rendered
  message: false
---

## Get dataset

```{r}
#| label: prepare-session
#| code-fold: true
library(ucimlrepo)
library(tidyverse)
library(gtsummary)
library(gt)

data<-fetch_ucirepo(id=890)
data1<- data$data
df<- data1$original
```

```{r}
#| label: tbl-summary-table
#| tbl-cap: "Summary of the characteristics of the patients included in the data"
#| code-fold: true

cdf<-df%>%
  select(str2, gender, age, race, homo, hemo, drugs, karnof, symptom, cd40, preanti)

#names(cdf)[2]<- "Male"

cdf$str2<- factor(cdf$str2, levels= c(0, 1),
                         labels= c("No Prior Exposure to Antiviral Agents", "Prior Exposure to Antiviral Agents"))
cdf$race<- factor(cdf$race, levels = c(0,1),
                  labels=c("White","Non-white"))
cdf<-cdf%>%mutate(karnof100=ifelse(karnof==100,"100%","Other"),preantimo=preanti/30.4)
Table1<-cdf %>%
  tbl_summary(include= c(str2,gender,age,race,homo,drugs,hemo,karnof100,symptom,cd40,preantimo),
              by=str2,
              label=list(gender~"Male sex",
                         age ~ "Age - yr",
                         race ~ "Race or ethnic group",
                         homo~"Homosexuality",
                         drugs~"Injection-drug use",
                         hemo~"Hemophilia",
                         karnof100~"Karnofsky score",
                         symptom~"Symptomatic HIV infection",
                         cd40~ "CD4 cell count -cells/mm^3",
                         preantimo~"Median length of prior antiviral therapy - mo (lower and upper quartiles)"),
              statistic = list(age~"{mean} ± {sd}",
                               cd40~"{mean} ± {sd}"),
              digits = list(age ~ c(1, 1)),
              missing="no")%>%
  add_overall()

Table1 |>
  modify_header(stat_0 ~ "**All Patients** N = {n}")|>
  modify_table_body(
    ~ .x |>
      dplyr::add_row(
        variable = "header_risk",
        label = "Risk factors",
        row_type = "label",
        .before = which(.x$variable == "homo")[1]
      )
  )|>
  modify_caption("**Table 1**")

```

@tbl-summary-table was reconstructed using the data available. From the onset none of the following analysis can 100% reconstruct the findings in the paper, simply because in the paper the list 2467 as the total patient number. However the data only includes 2193 patient. Meaning we have at least: `r 2467-2193` data points less to include in the analysis. However it is possible that the data we are working with has already excluded any patients with recorded NAs.
