---
title: "Progression of Disease section"
author: "Natallia"
date: today
toc: true
format: 
  html:
    embed-resources: true
    code-fold: true
    # number-sections: true
    
editor: visual
editor-options:
  chunk-output-type: console
  
project:
  output-dir: output
  execute-dir: project

execute: 
  warning: false
  error: false
  freeze: true
---

Definition of the variables:
- treatment indicator (0 = ZDV only; 1 = ZDV + ddI, 2 = ZDV + Zal, 3 = ddI only)
- hemo hemophilia (0=no, 1=yes)
- drugs history of IV drug use (0=no, 1=yes)
- oprior Non-ZDV antiretroviral therapy pre-175 (0=no, 1=yes)

```{r}
#| label: preparing-session

# Load the package
library(ucimlrepo)

# Fetch the dataset by its UCI ID (890)
aids_clinical_trials_group_study_175 <- fetch_ucirepo(id = 890)

# View metadata
# print(aids_clinical_trials_group_study_175$metadata)

# Access features and targets
features <- aids_clinical_trials_group_study_175$data$features
targets  <- aids_clinical_trials_group_study_175$data$targets

features$cid<-targets$cid

```

```{r}
#| label: fig-time-dist

hist(features$time/7)
quantile(features$time/7) #143 is the median treatment weeks

quantile(features$time)
#   0%  25%  50%  75% 100% 
#   14  727  997 1091 1231 
# > 1231/365
# [1] 3.372603
# > 997/365
# [1] 2.731507 => average 2.7 years treatment

#str2 antiretroviral history (0=naive, 1=experienced)
features$str2<-factor(features$str2, levels=c(0,1), labels=c("naive", "experienced"))
tapply(features$time/7, features$str2, median)
```


::: {.callout-note}
Death can not be determined. As well as the cd40 counts at baseline include <200 cells per mm3 which doesnt make sense!!
:::

```{r}
library(gtsummary)

#counts declide -> cd40 and cd420 ratio
#AIDS -> ower than 200 cd420
#death -> karnof score 0 !!!! => no karnof == 0


table(features$cd420/features$cd40<=0.5 | features$cd420 < 200)
table(features$cd420<200)
table(features$karnof==0)
summary(features$karnof)


# features$event<-ifelse(
#   # features$karnof==0 |
#   features$cd420/features$cd40<=0.5 |
#     features$cd420 < 200, 1, 0
# )

```
```{r}
# features$trt<-factor(features$trt, labels=c("ZDV", "ZDV + ddI", "ZDV + Zal", "Didanosine"))
# 
# surv_obj <- Surv(time = features$time/7, event = features$event)
# km_fit <- survfit(surv_obj ~ trt, data = features)
# 
# 
# plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
#      xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
# legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
#        col = c("blue", "red", "green", "violet"), lty = 1:2)
# 
# 
# features$event_sec<-ifelse(
#   # features$karnof==0 |
#     features$cd420 < 200, 1, 0
# )
# 
# surv_obj_2 <- Surv(time = features$time/7, event = features$event_sec)
# km_fit_2 <- survfit(surv_obj_2 ~ trt, data = features)
# 
# 
# plot(km_fit_2, col = c("blue", "red", "green", "violet"), lty = 1:2,
#      xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
# legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
#        col = c("blue", "red", "green", "violet"), lty = 1:2)
# 


# features%>%
#   tbl_summary( 
#     by=trt,
#     include= c()
#       )


```

# 1st Paragraph
Five hundred sixty-five patients reached a primary
end point — a 50 percent decline in the CD4 cell
count, an AIDS-defining event, or death. Sixty
```{r}
prim<-table(features$cid)

#of these how many in prior antiretroviral therapy (1,0)
features$str2<-factor(features$str2, levels=c(0,1), labels=c("no", "yes"))
prim_str2<-table(features$cid, features$str2)

prim_str2[2,1]/prim[2] #NO antiretroviral
prim_str2[2,2]/prim[2] #antiretroviral

```
We got ` r prim[2]` people who reached the primary endpoint.


# 2nd paragraph
"Sixty-nine
percent of these patients were in the subgroup with
prior antiretroviral therapy, and 31 percent in the
subgroup with no prior antiretroviral therapy."

Just as the paper described we've got similar results considering the percentages of people with/or without prior antiretroviral therapy out of the ones who reached the primary endpoint: `r prim_str2[2,2]/prim[2]*100`% btw `r prim_str2[2,1]/prim[2]*100`%.

# 3rd paragraph
End points related to the CD4 cell count constituted the
majority of the first events (399 of 565, or 71 percent).

```{r}
table(features$cid, features$cd420/features$cd40<=0.5 | features$cd420<200)

135/prim[2]
# 0.2591171 
```

=> In our analysis we could only find `r 135/prim[2]*100`% of the primary endpoints related to CD4+ cells in both vases of decrease more than 50% and the AIDS phase with CD4+<200 cells per mm3.


# 4th paragraph
In 75 percent of cases the primary end point
was reached during treatment or within three months
of its discontinuation; in 15 percent of cases the end
point was reached more than one year after treatment
was discontinued. Of the clinical (secondary) end
points of an AIDS-defining event or death or death
alone, 50 and 25 percent, respectively, occurred during
treatment or within three months after its discontinuation;
25 and 35 percent, respectively, occurred
more than one year after its discontinuation.

:::{.callout-note}
Impossible to determine this percentages -> only 1 time variable, we can not see the follow-up or discontinuation of the treatment point
:::


# Figures reproduced: Figure 1 A, B

```{r}

library(survival)

surv_obj <- Surv(time = features$time/7, event = features$cid)
km_fit <- survfit(surv_obj ~ trt, data = features)


plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)


features$sec_endpoint_maybe<-ifelse(features$cid==1 & features$cd420/features$cd40>0.5,1,0)


surv_obj <- Surv(time = features$time/7, event = features$sec_endpoint_maybe)
km_fit <- survfit(surv_obj ~ trt, data = features)


plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)

```

