---
title: "Progression of Disease section"
author: "Natallia"
date: today
toc: true
format: 
  html:
    embed-resources: true
    code-fold: true
    # number-sections: true
    
editor: visual
editor-options:
  chunk-output-type: console
  
project:
  output-dir: output
  execute-dir: project

execute: 
  warning: false
  error: false
  freeze: true
---

Definition of the variables: - treatment indicator (0 = ZDV only; 1 = ZDV + ddI, 2 = ZDV + Zal, 3 = ddI only) - hemo hemophilia (0=no, 1=yes) - drugs history of IV drug use (0=no, 1=yes) - oprior Non-ZDV antiretroviral therapy pre-175 (0=no, 1=yes)

```{r}
#| label: preparing-session
#| echo: false

# Load the package
library(ucimlrepo)

# Fetch the dataset by its UCI ID (890)
aids_clinical_trials_group_study_175 <- fetch_ucirepo(id = 890)

# View metadata
# print(aids_clinical_trials_group_study_175$metadata)

# Access features and targets
features <- aids_clinical_trials_group_study_175$data$features
targets  <- aids_clinical_trials_group_study_175$data$targets

features$cid<-targets$cid
features$trt<-factor(features$trt,
                     levels=c(0,1,2,3),
                     labels=c("Zidovudine", "Zidovudine and Didanosine", "Zidovudine and Zalcitabine", "Didanosine"))

```

```{r}
#| label: fig-time-dist

# hist(features$time/7)
# quantile(features$time/7) #143 is the median treatment weeks
# 
# quantile(features$time)
#   0%  25%  50%  75% 100% 
#   14  727  997 1091 1231 
# > 1231/365
# [1] 3.372603
# > 997/365
# [1] 2.731507 => average 2.7 years treatment

#str2 antiretroviral history (0=naive, 1=experienced)
# features$str2<-factor(features$str2, levels=c(0,1), labels=c("naive", "experienced"))
# tapply(features$time/7, features$str2, median)

```

```{r}
#| label: fig-cd40-baseline
#| fig-cap: "Baseline distribution of CD4+ cell counts per mm^3"
#| echo: false
quantile(features$cd40)
cd40_200<-table(features$cd40<200)

h<-hist(features$cd40, plot=FALSE)
cols <- ifelse(h$mids < 200, "red", "gray")


hist(features$cd40,
     main="",
     ylab="Frequency",
     xlab="CD4+ cell counts per mm^3",
     col=cols)
```

As we can see from the histogram there appear to be people with less than 200 CD4+ cells per mm\^3 which would be against the criteria mentioned in the paper. (`r round(cd40_200[2]/nrow(features)*100, 3)` %)

::: callout-note
Death can not be determined. As well as the cd40 counts at baseline include \<200 cells per mm3 which does not make sense!!
:::

```{r}
#| include: false
library(gtsummary)

#counts declide -> cd40 and cd420 ratio
#AIDS -> ower than 200 cd420
#death -> karnof score 0 !!!! => no karnof == 0


table(features$cd420/features$cd40<=0.5 | features$cd420 < 200)
table(features$cd420<200)
table(features$karnof==0)
summary(features$karnof)


# features$event<-ifelse(
#   # features$karnof==0 |
#   features$cd420/features$cd40<=0.5 |
#     features$cd420 < 200, 1, 0
# )

```

# 1st Paragraph

*Five hundred sixty-five patients reached a primary end point — a 50 percent decline in the CD4 cell count, an AIDS-defining event, or death.*

```{r}
#| label: pr-1
#| output: false
prim<-table(features$cid)

#of these how many in prior antiretroviral therapy (1,0)
features$str2<-factor(features$str2, levels=c(0,1), labels=c("no", "yes"))
prim_str2<-table(features$cid, features$str2)

prim_str2[2,1]/prim[2] #NO antiretroviral
prim_str2[2,2]/prim[2] #antiretroviral

```

We got `r prim[2]` people who reached the primary endpoint. (`r round(prim[2]/nrow(features)*100, 3)` %)

# 2nd paragraph

*"Sixty-nine percent of these patients were in the subgroup with prior antiretroviral therapy, and 31 percent in the subgroup with no prior antiretroviral therapy."*

Just as the paper described we've got similar results considering the percentages of people with/or without prior antiretroviral therapy out of the ones who reached the primary endpoint: `r round(prim_str2[2,2]/prim[2]*100,3)`% btw `r round(prim_str2[2,1]/prim[2]*100,3)`%.

# 3rd paragraph

*End points related to the CD4 cell count constituted the majority of the first events (399 of 565, or 71 percent).*

```{r}
#| output: false
(table(features$cid, features$cd420/features$cd40<=0.5|features$cd420<200))

135/prim[2]
# 0.2591171 
```

In our analysis we could only find `r round(135/prim[2]*100,3)`% of the primary endpoints related to CD4+ cells in both vases of decrease more than 50% and the AIDS phase with CD4+\<200 cells per mm3.

# 4th paragraph

*In 75 percent of cases the primary end point was reached during treatment or within three months of its discontinuation; in 15 percent of cases the end point was reached more than one year after treatment was discontinued. Of the clinical (secondary) end points of an AIDS-defining event or death or death alone, 50 and 25 percent, respectively, occurred during treatment or within three months after its discontinuation; 25 and 35 percent, respectively, occurred more than one year after its discontinuation.*

::: callout-note
Impossible to determine this percentages -\> only 1 time variable, we can not see the follow-up or discontinuation of the treatment point
:::

# Figures reproduced: Figure 1 A, B

```{r}
#| label: fig-1ab-reproduced
#| fig-cap: "Kaplan-Meier Estimates of the Proportion of Patients Not Reaching the Primary Endpoint (A) or the Secondary Endpoint(B)"
#| fig-width: !expr 30/2.54
#| fig-height: !expr 15/2.54

#1. Kaplan Meier
library(survival)

surv_obj1 <- Surv(time = features$time/7, event = features$cid)
km_fit1 <- survfit(surv_obj1 ~ trt, data = features)



features$sec_endpoint_maybe<-ifelse(features$cid==1 & features$cd420/features$cd40>0.5,1,0)


surv_obj2 <- Surv(time = features$time/7, event = features$sec_endpoint_maybe)
km_fit2 <- survfit(surv_obj2 ~ trt, data = features)


par(mfrow=c(1,2))
plot(km_fit1, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)


plot(km_fit2, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)


```




```{r}
# 2. log-rank test
log_rank_1<-survdiff(Surv(time/7, cid) ~trt, features)
log_rank_2<-survdiff(Surv(time/7, sec_endpoint_maybe) ~trt, features)

```
Log Rank test looks if there is a difference between the KM curves. Null-Hypothesis says there is no
difference and thus we want to disprove it.
Use Chi squared as we have count/categorical data. (For continuos data we use F test.)


```{r}
# 3. Cox proportionality hazard model stratified by strat
cox_1<-coxph(Surv(time/7, cid)~trt + strata(strat), features)
cox_2<-coxph(Surv(time/7, sec_endpoint_maybe)~trt + strata(strat), features)

```
It models how different variables affect the risk of an event over time.
In model 1 we can see that treatment 1 (ZZI + ddI) has the lowest risk to develop the event CID of hazard ratio of `r round(exp(cox_1$coefficients[1]),3)` thus a `r (1-round(exp(cox_1$coefficients[1]),3))*100`% lower risk.
For model 2 with the secondary endpoint we see that also treatment 1 has the lowest risk. (`r round(exp(cox_2$coefficients[1]),3)` thus `r (1-round(exp(cox_2$coefficients[1]),3))*100`% lower risk)


```{r}
# 4. Cox "" with interaction term for diff strat
cox_inter_1<-coxph(Surv(time/7, cid)~trt * strat, features)
cox_inter_2<-coxph(Surv(time/7, sec_endpoint_maybe)~trt *strat, features)
```
The interaction term is not significant as shown in the Cox proportionality hazard model. Thus we can see that the treatment works the same in each stratification group.


Treatment 1 showed significant protective effect in Kaplan Meier, log-rank test and simple Cox models, but was not
statistically significant in the interaction model, likely due to reduced precision from additional parameters.
Overall, results show a consistent protective effect of treatment 1.

# Reconstructing the Table 2 with analysis results
```{r}
#| label: tbl-1
#| tbl-cap: "Rates of Disease progression"

library(gtsummary)

#compute N and (%)
library(dplyr)
n_trt<- features %>%
  filter(cid==1)%>%
  count(trt) %>%
  mutate(
    pct=n/sum(n)*100,
    n_pct = sprintf("%d (%.0f)", n, pct)
  )

#compute regression tables

tbl_1<-tbl_regression(cox_1, exponentiate = T,
  label = list(trt ~ "Treatment"))%>%
  add_global_p()%>%
  modify_header(label ~ "")


tbl_2<-tbl_regression(cox_2, exponentiate = T,
  label = list(trt ~ "Treatment"))%>%
  add_global_p()%>%
  modify_header(label ~ "")%>%
    modify_table_styling(
    columns = label,
    align = "left",
  ) %>%
  modify_table_body(
    ~ .x %>%
      left_join(n_trt, by = c("label" = "trt"))
  )

tbl_2 <-
  tbl_2 %>%
  modify_header(n_pct ~ "N (%)")


tbl_merge(
  tbls = list(tbl_1, tbl_2),
  tab_spanner = c("Primary Endpoint", "Secondary Endpoint")
)
```

