---
title: "Progression of Disease section"
author: "Natallia"
date: today
toc: true
format: 
  html:
    embed-resources: true
    code-fold: true
    number-sections: true
    
editor: visual
editor-options:
  chunk-output-type: console
  
project:
  output-dir: output
  execute-dir: project

execute: 
  warning: false
  error: false
  freeze: true
---

Definition of the variables:
- treatment indicator (0 = ZDV only; 1 = ZDV + ddI, 2 = ZDV + Zal, 3 = ddI only)
- hemo hemophilia (0=no, 1=yes)
- drugs history of IV drug use (0=no, 1=yes)
- oprior Non-ZDV antiretroviral therapy pre-175 (0=no, 1=yes)

```{r}
#| label: preparing-session

# Load the package
library(ucimlrepo)

# Fetch the dataset by its UCI ID (890)
aids_clinical_trials_group_study_175 <- fetch_ucirepo(id = 890)

# View metadata
# print(aids_clinical_trials_group_study_175$metadata)

# Access features and targets
features <- aids_clinical_trials_group_study_175$data$features
targets  <- aids_clinical_trials_group_study_175$data$targets

features$cid<-targets$cid

```

```{r}
#| label: fig-time-dist

hist(features$time/7)
quantile(features$time/7) #143 is the median

#str2 antiretroviral history (0=naive, 1=experienced)
features$str2<-factor(features$str2, levels=c(0,1), labels=c("naive", "experienced"))
tapply(features$time/7, features$str2, median)
```


::: {.callout-note}
Death can not be determined. As well as the cd40 counts at baseline include <200 cells per mm3 which doesnt make sense!!
:::

```{r}
library(gtsummary)

#counts declide -> cd40 and cd420 ratio
#AIDS -> ower than 200 cd420
#death -> karnof score 0 !!!! => no karnof == 0


table(features$cd420/features$cd40<=0.5 | features$cd420 < 200)
table(features$cd420<200)
table(features$karnof==0)
summary(features$karnof)


# features$event<-ifelse(
#   # features$karnof==0 |
#   features$cd420/features$cd40<=0.5 |
#     features$cd420 < 200, 1, 0
# )

```
```{r}
# features$trt<-factor(features$trt, labels=c("ZDV", "ZDV + ddI", "ZDV + Zal", "Didanosine"))
# 
# surv_obj <- Surv(time = features$time/7, event = features$event)
# km_fit <- survfit(surv_obj ~ trt, data = features)
# 
# 
# plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
#      xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
# legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
#        col = c("blue", "red", "green", "violet"), lty = 1:2)
# 
# 
# features$event_sec<-ifelse(
#   # features$karnof==0 |
#     features$cd420 < 200, 1, 0
# )
# 
# surv_obj_2 <- Surv(time = features$time/7, event = features$event_sec)
# km_fit_2 <- survfit(surv_obj_2 ~ trt, data = features)
# 
# 
# plot(km_fit_2, col = c("blue", "red", "green", "violet"), lty = 1:2,
#      xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
# legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
#        col = c("blue", "red", "green", "violet"), lty = 1:2)
# 


# features%>%
#   tbl_summary( 
#     by=trt,
#     include= c()
#       )


```

```{r}

library(survival)

surv_obj <- Surv(time = features$time/7, event = features$cid)
km_fit <- survfit(surv_obj ~ trt, data = features)


plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)


features$sec_endpoint_maybe<-ifelse(features$cid==1 & features$cd420/features$cd40>0.5,1,0)


surv_obj <- Surv(time = features$time/7, event = features$sec_endpoint_maybe)
km_fit <- survfit(surv_obj ~ trt, data = features)


plot(km_fit, col = c("blue", "red", "green", "violet"), lty = 1:2,
     xlab = "Time", ylab = "Survival probability", ylim = c(0,1))
legend("bottomleft", legend = c("ZDV"," ZDV + ddI"," ZDV + Zal"," Didanosine"),
       col = c("blue", "red", "green", "violet"), lty = 1:2)

```

